---
title: "Heterosis effect"
---


```{r}
library("glmmTMB")
library("dplyr")
library("DHARMa")
library("ggplot2")
library("ggpubr")
library("tidyverse")
library("lme4")
library("emmeans")
```

parameters for plotting

```{r}
my_theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5))
```


Read data

```{r}
Dataset <- read.csv(here::here("dataset.csv"))

# Set random effects as factors

Dataset$block <- as.factor(Dataset$block)
Dataset$cross <- as.factor(Dataset$cross)

Dataset$bushiness[Dataset$bushiness==0] <- NA

# to test heterosis, we discard hybrids

Dataset <-subset(Dataset, Dataset$colors != "M_Y") 

Dataset <- Dataset |> 
  rename(heterosis = w_b)

head(Dataset)
```

Create a dataset excluding control to run the survival model:

```{r}
Dataset_nc<-subset(Dataset, Dataset$trt != "control") #Dataset used for the model with survival as response variable
```

```{r}
glmer_surv_nc <- glmmTMB(survived~ trt + heterosis + trt:heterosis + (1|block) +(1|cross), family = "binomial", data = 
                         Dataset_nc)

glmer_f <- glmmTMB(flowering~trt + heterosis + trt:heterosis + (1|block) +(1|cross), family = "binomial", data = Dataset)

glmer_nf_d <- glmmTMB(flower_count ~ trt + heterosis + trt:heterosis, data = Dataset, family = "nbinom2")


lmer_height<-lmer(height_cm~trt + heterosis + trt:heterosis + (1|block) +(1|cross), Dataset, na.action=na.exclude)
lmer_height.o<-lmer(height_cm~trt + heterosis + trt:heterosis + (1|block)+(1|cross), Dataset, na.action=na.omit) # This model only changes the na.action because packages used for diagnostics cant handle na.exclude() and some plotting functions handle na.omit() better. This does not affect any estimate and both models are equivalent as both na.exclude and na.omit remove the row where any of the dependent or independent variable is missing

glmer_bush <- glmmTMB(bushiness~ trt + heterosis + trt:heterosis + (1|block)+(1|cross), data = Dataset, family = "poisson")


glmer_tf <- glmmTMB(days_1flower~trt + heterosis + trt:heterosis + (1|block) +(1|cross), family = "nbinom2", data = Dataset)
```


## Model evaluations

```{r}

# Survival

simulationOutput_surv_d<-simulateResiduals(glmer_surv_nc) 
dharma_plot_surv<-plot(simulationOutput_surv_d)
surv_dharma<-testResiduals(simulationOutput_surv_d)


# Flowering

simulationOutput_flow<-simulateResiduals(glmer_f) 
dharma_plot_flow<-plot(simulationOutput_flow)
flow_dharma<-testResiduals(simulationOutput_flow) 


# flower_count


simulationOutput_nflow<-simulateResiduals(glmer_nf_d) 
dharma_plot_nflow<-plot(simulationOutput_nflow)



# Height

simulationOutput_height_d<-simulateResiduals(lmer_height.o)
dharma_plot_height<-plot(simulationOutput_height_d)
height_dharma<-testResiduals(simulationOutput_height_d)


# Bushiness

simulationOutput_bush_d<-simulateResiduals(glmer_bush)
dharma_plot_bush<-plot(simulationOutput_bush_d)
bush_dharma<-testResiduals(simulationOutput_bush_d) 


# t flower

simulationOutput_tflow_d<-simulateResiduals(glmer_tf)
dharma_plot_tflow<-plot(simulationOutput_tflow_d)
tflow_dharma<-testResiduals(simulationOutput_tflow_d)


```


## Summary Anova tables

```{r}

s_surv<-car::Anova(glmer_surv_nc,type="II")
s_flow<-car::Anova(glmer_f,type="II")
s_nflow<-car::Anova(glmer_nf_d,type="II")
s_height<-car::Anova(lmer_height,type="III")
s_bush<-car::Anova(glmer_bush,type="III")
s_tflow<-car::Anova(glmer_tf,type="III")
#s_iheight<-car::Anova(lm_iheight_col,type="III")

options(scipen = 999)

surv<-as.data.frame(s_surv)
colnames(surv)[1] <- "Chisq"
surv <- cbind(surv,rep("survival", nrow(surv)))
colnames(surv)[ncol(surv)] = "trait"

flow<-as.data.frame(s_flow)
colnames(flow)[1] <- "Chisq"
flow <- cbind(flow,rep("flowering", nrow(flow)))
colnames(flow)[ncol(flow)] = "trait"


nflow<-as.data.frame(s_nflow)
nflow <- cbind(nflow,rep("n flowers", nrow(nflow)))
colnames(nflow)[ncol(nflow)] = "trait"


height<-as.data.frame(s_height)
height <- cbind(height,rep("height", nrow(height)))
colnames(height)[ncol(height)] = "trait"


bush<-as.data.frame(s_bush)
bush <- cbind(bush,rep("bushiness", nrow(bush)))
colnames(bush)[ncol(bush)] = "trait"


tflow<-as.data.frame(s_tflow)
colnames(tflow)[1] <- "Chisq"
tflow <- cbind(tflow,rep("time to flower", nrow(tflow)))
colnames(tflow)[ncol(tflow)] = "trait"


#iheight<-as.data.frame(s_iheight)

df_anovas <- rbind(surv,
      flow,
      nflow,
      tflow,
      height,
      bush)

df_anovas$"Pr(>Chisq)"<- round(df_anovas$"Pr(>Chisq)", 4)
df_anovas$"Chisq"<- round(df_anovas$"Chisq", 2)

df_anovas

#write.csv(df_anovas, here::here("tables/df_anovas_heterosis.csv"))

```


## Group comparisons of predicted means & plotting


Plotting settings

```{r}

pd = position_dodge(0.9)
alpha_raw_points = 0.15

trt_order <- c("control", "early drought", "late drought")

Dataset$trt <- factor(Dataset$trt,
    levels = trt_order,ordered = TRUE)

```


### ----- Survival - all levels

Compute emmeans and CI

```{r}

marginal_surv1.1 = emmeans(glmer_surv_nc,
                   ~ heterosis:trt,  calc = c(n = ".wgt."))

back.emmeans_surv<-confint(marginal_surv1.1, type = "response")

```
Add letter display from tukey tests

```{r}


surv_emmeans_cld <- multcomp::cld(object = marginal_surv1.1,
                           Letters = letters)


surv_emmeans_cld <- surv_emmeans_cld %>% 
  arrange(heterosis) %>% 
  select(trt, .group, heterosis)


back.emmeans_surv <- left_join(back.emmeans_surv, surv_emmeans_cld, by = c("trt","heterosis"))

back.emmeans_surv

```


```{r}


Dataset_percentage <- aggregate(survived ~ trt + heterosis, data = Dataset, FUN = mean)

Dataset_percentage$trt <- factor(Dataset_percentage$trt,
   levels = trt_order,ordered = TRUE)



# Plotting raw data points
surv_raw <- ggdotchart(
  Dataset_percentage, 
  x = "trt", 
  y = "survived",
  color = "heterosis", 
  palette = c("black", "black"), 
  alpha = 0.3,
  size = 5, 
  shape = 21,
  #add = "segment", 
  #add.params = list(color = "lightgray", size = 1.3),
  position = position_dodge(0.6),
  ggtheme = theme_classic(),
  sorting = "none",
  order = trt_order
)

# Overlay emmeans and CI on the raw data plot
surv_overlay_plot <- surv_raw +
  geom_errorbar(
    data = back.emmeans_surv,
    aes(
      x = trt, #factor(trt, levels = trt_order), 
      y = prob,  # Assuming prob is the probability in decimal form
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      color = heterosis
    ),
    position = position_dodge(width = 0.6),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_surv,
    aes(
      x = trt, #factor(trt, levels = trt_order), 
      y = prob,  # Assuming prob is the probability in decimal form
      color = heterosis
    ),
    position = position_dodge(width = 0.6),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
 #geom_text(
  #  data = back.emmeans_surv,
  #  aes(
  #    x = trt, # Adjust position based on factor levels,
  #    color = heterosis,
  #    y = asymp.UCL + 0.02,  # Adjust the position of the letters
 #     label = .group  # Use the .group column for labels
 #   ),
 #   position = position_dodge(width = 0.6),
 #   vjust = -0.5,  # Adjust vertical position
 #   size = 4
 # ) +
  scale_fill_manual(values = c("red", "blue")) +
  scale_color_manual(
    values = c("red", "blue"),  # Remove black color
    labels = c("between populations", "within populations")  # Update labels
  ) +
  ylim(0,1.05)+
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  labs(color = "cross") +
#  theme(legend.position = "none") +
  xlab(NULL) +
  ylab("survival probability") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))


surv_overlay_plot





```


### ----- Flowering - All levels


Compute emmeans and CI

```{r}
emmeans_flowering = emmeans(glmer_f,
                   ~ heterosis:trt,  calc = c(n = ".wgt."))

back.emmeans_flowering<-confint(emmeans_flowering, type = "response")

```

Add letter display from tukey tests

```{r}


flow_emmeans_cld <- multcomp::cld(object = emmeans_flowering,
                           Letters = letters)

flow_emmeans_cld <- flow_emmeans_cld %>% 
  arrange(trt, heterosis) %>% 
  select(.group, heterosis, trt)


back.emmeans_flowering <- left_join(back.emmeans_flowering, flow_emmeans_cld, by = c("trt", "heterosis"))

back.emmeans_flowering

```


```{r}

Dataset_percentage <- aggregate(flowering ~ trt + heterosis, data = Dataset, FUN = mean)


# Plotting raw data points
flow_raw <- ggdotchart(
  Dataset_percentage, 
  x = "trt", 
  y = "flowering",
  color = "heterosis", 
  palette = c("black", "black", "black"), 
  alpha = 0.3,
  size = 5, 
  shape = 21,
  #add = "segment", 
  #add.params = list(color = "lightgray", size = 1.3),
  position = position_dodge(0.6),
  ggtheme = theme_classic(),
    sorting = "none",
  order = trt_order
)

# Overlay emmeans and CI on the raw data plot
flow_overlay_plot <- flow_raw +
  geom_errorbar(
    data = back.emmeans_flowering,
    aes(
      x = trt, #factor(trt, levels = c(trt_order)), 
      y = prob,  # Assuming prob is the probability in decimal form
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      color = heterosis
    ),
    position = position_dodge(width = 0.6),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_flowering,
    aes(
      x = trt, #factor(trt, levels = c(trt_order)), 
      y = prob,  # Assuming prob is the probability in decimal form
      color = heterosis
    ),
    position = position_dodge(width = 0.6),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
 # geom_text(
 #   data = back.emmeans_flowering,
 #   aes(
 #     x = trt, # Adjust position based on factor levels,
  #    color = heterosis,
  #    y = asymp.UCL + 0.02,  # Adjust the position of the letters
  #    label = .group  # Use the .group column for labels
 #   ),
 #   position = position_dodge(width = 0.6),
 #   vjust = -0.5,  # Adjust vertical position
 #   size = 4
#  ) +
 scale_fill_manual(values = c("red", "blue")) +
  scale_color_manual(
    values = c("red", "blue"),  # Remove black color
    labels = c("between populations", "within populations")  # Update labels
  ) +
  ylim(0, 1.05)+
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  labs(color = "cross") +
 # theme(legend.position = "none")+
  xlab(NULL) +
  ylab("flowering prob. (all plants)") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))


flow_overlay_plot

```



### ----- N flowers - All levels

Compute emmeans and CI


```{r}
### N flowers (1 drought level) - treatment significant

marginal_nf.d = emmeans(glmer_nf_d,
                   ~ heterosis:trt,  calc = c(n = ".wgt."))

back.emmeans_nflow<-confint(marginal_nf.d, type = "response")


back.emmeans_nflow[which(is.infinite(back.emmeans_nflow$asymp.UCL)),"asymp.UCL"] <- 0

```



Add letter display from tukey tests

```{r}


nflow_emmeans_cld <- multcomp::cld(object = marginal_nf.d,
                           Letters = letters)

nflow_emmeans_cld <- nflow_emmeans_cld %>% 
  arrange(trt, heterosis) %>% 
  select(.group, heterosis, trt)


back.emmeans_nflow <- left_join(back.emmeans_nflow, nflow_emmeans_cld, by = c("trt", "heterosis"))

back.emmeans_nflow

```





```{r}


nflow_raw <- ggplot(Dataset, aes(x = trt, y = flower_count, group = heterosis)) +
  geom_jitter(
    aes(color = "black", fill = "black"),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  scale_fill_manual(values=c("black", "black"))+
  scale_color_manual(values=c("black", "black"))+
  theme_classic() +
  my_theme




# Overlay emmeans and CI on the raw data plot
nflow_overlay_plot <- ggplot(Dataset, aes(x = trt, y = flower_count, group = heterosis)) +
  geom_jitter(
    aes(color = heterosis, fill = heterosis),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  geom_errorbar(
    data = back.emmeans_nflow,
    aes(x = trt, y = response, ymin = asymp.LCL, ymax = asymp.UCL, color = heterosis),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_nflow,
    aes(x = trt, y = response, color = heterosis),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
  # Define the scales only once
  scale_color_manual(
    values = c("red", "blue"),
    labels = c("between populations", "within populations")
  ) +
  scale_fill_manual(
    values = c("red", "blue")
  ) +
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  labs(color = "cross") +
  xlab(NULL) +
  ylab("flower count (survivors)") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))

nflow_overlay_plot


```



### ----- Height


```{r}

emmeans_height = emmeans(lmer_height,
                   ~ heterosis:trt,  calc = c(n = ".wgt."))

emmeans_height <- multcomp::cld(object = emmeans_height,
                                   Letters = letters)

```




RAW DATA



```{r}


# Define the base plot with jitter for raw data
height_raw <- ggplot(Dataset, aes(x = trt, y = height_cm, group = heterosis)) +
  geom_jitter(
    aes(color = heterosis, fill = heterosis),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  theme_classic() +
  my_theme

# Overlay emmeans and CI on the raw data plot
height_overlay_plot <- height_raw +
  geom_errorbar(
    data = emmeans_height,
    aes(x = trt, y = emmean, ymin = lower.CL, ymax = upper.CL, color = heterosis),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = emmeans_height,
    aes(x = trt, y = emmean, color = heterosis),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
  # Define the scales only once
  scale_color_manual(
    values = c("red", "blue"),
    labels = c("between populations", "within populations")
  ) +
  scale_fill_manual(
    values = c("red", "blue")
  ) +
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  labs(color = "cross") +
  xlab(NULL) +
  ylab("height (cm)") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))

height_overlay_plot


```


### ----- Bushiness

Compute emmeans and CI

```{r}
emmeans_bush = emmeans(glmer_bush,
                   ~ heterosis:trt,  calc = c(n = ".wgt."))

back.emmeans_bush<-confint(emmeans_bush, type = "response")


```


Add letter display from tukey tests

```{r}

emmeans_bush_cld <- multcomp::cld(object = emmeans_bush,
                                   Letters = letters)

emmeans_bush_cld <- emmeans_bush_cld %>% 
  arrange(trt, heterosis) %>% 
  select(.group, heterosis, trt)


back.emmeans_bush <- left_join(back.emmeans_bush, emmeans_bush_cld, by = c("trt", "heterosis"))

back.emmeans_bush

```




```{r}


# Define the base plot with jitter for raw data
bush_raw <- ggplot(Dataset, aes(x = trt, y = bushiness, group = heterosis)) +
  geom_jitter(
    aes(color = heterosis, fill = heterosis),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  theme_classic() +
  my_theme

# Overlay emmeans and CI on the raw data plot
bushiness_overlay_plot <- bush_raw +
  geom_errorbar(
    data = back.emmeans_bush,
    aes(x = trt, y = rate, ymin = asymp.LCL, ymax = asymp.UCL, color = heterosis),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_bush,
    aes(x = trt, y = rate, color = heterosis),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
  # Define the scales only once
  scale_color_manual(
    values = c("red", "blue"),
    labels = c("between populations", "within populations")
  ) +
  scale_fill_manual(
    values = c("red", "blue")
  ) +
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  labs(color = "cross") +
  xlab(NULL) +
  ylab("bushiness (n nodes)") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))

bushiness_overlay_plot



```


### ----- Time to flower


Compute emmeans and CI

```{r}
emmeans_tflow = emmeans(glmer_tf,
                   ~ heterosis:trt,  calc = c(n = ".wgt."))

back.emmeans_tflow<-confint(emmeans_tflow, type = "response")

```


Add letter display from tukey tests

```{r}

emmeans_tflow_cld <- multcomp::cld(object = emmeans_tflow,
                                   Letters = letters)

emmeans_tflow_cld <- emmeans_tflow_cld %>% 
  arrange(trt, heterosis) %>% 
  select(.group, heterosis, trt)


back.emmeans_tflow <- left_join(back.emmeans_tflow, emmeans_tflow_cld, by = c("trt", "heterosis"))

back.emmeans_tflow

```






```{r}
# Define the base plot with jitter for raw data
tflow_raw <- ggplot(Dataset, aes(x = trt, y = days_1flower, group = heterosis)) +
  geom_jitter(
    aes(color = heterosis, fill = heterosis),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  theme_classic() +
  my_theme

# Overlay emmeans and CI on the raw data plot
tflow_overlay_plot <- tflow_raw +
  geom_errorbar(
    data = back.emmeans_tflow,
    aes(x = trt, y = response, ymin = asymp.LCL, ymax = asymp.UCL, color = heterosis),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_tflow,
    aes(x = trt, y = response, color = heterosis),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
  # Define the scales only once
  scale_color_manual(
    values = c("red", "blue"),
    labels = c("between populations", "within populations")
  ) +
  scale_fill_manual(
    values = c("red", "blue")
  ) +
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  labs(color = "cross") +
 # ylim(38, 95) +
  xlab(NULL) +
  ylab("time to flower (days)") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))

tflow_overlay_plot


```


# Fig heterosis

```{r}


fig_heterosis <- ggarrange(
  surv_overlay_plot,
  flow_overlay_plot,
  nflow_overlay_plot,
  height_overlay_plot,
  bushiness_overlay_plot,
  tflow_overlay_plot,
  
  ncol = 2,
  nrow = 3,
  
  labels = LETTERS[1:6],
  
  common.legend = TRUE,
  
  legend = "top"
)

fig_heterosis

#ggsave(here::here("figures/fig.heterosis.png"), height = 10, width = 10)


```

