---
title: "Mixed models - flower color and treatmenr"
output: html_notebook
---

This script conducts analyses to test the effect of treatment (control, early drought and late drought) and color or phenotype (A.m. pseudomajus, F1 hybrids, and A.m. striatum) on individual trait responses.



Load packages

```{r, results="hide", message=FALSE, waring = FALSE}


## load required packages
pkgs <- c("lme4", "arm", "multcomp", "nlme", 
          "lsmeans", "tidyverse","nlme", "tidyverse", 
          "ResourceSelection", "dplyr","effsize","hrbrthemes", 
          "viridis", "plyr", "cowplot", "aods3", 
          "AER", "pscl","DHARMa", "plotly", "ggpubr", 
          "extrafont","car", "rcompanion", "agricolae",
          "jtools", "sjPlot", "VGAM", "pracma", "rgl",
          "MASS", "performance", 
         # "brms", 
          "sjstats", 
          "sjmisc",
         #"lmerTest",
         #"rockchalk",
          "emmeans", 
         #"afex",
         #"effects",
         "forcats", 
          "visreg", "rstatix", 
         #"lmerTest",
          #"ggstance", 
         #"broomExtra",
         "dplyr", "grid",
         "glmmTMB",
         "ggeffects",
         "patchwork")


lapply(pkgs, library, character.only = TRUE)

rm(pkgs)
```


### Read data

the dataset used for the analyses is computed in **"compute_df_analyses.Rmd"**.

```{r}

Dataset <- read.csv("../dataset.csv")

# Filter to conduct analysis only for individuals resulted from between population crosses (discard within pop crosses)

Dataset <- Dataset %>% 
  filter(w_b == "b") 

# Set random effects as factors

Dataset$block <- as.factor(Dataset$block)
Dataset$cross <- as.factor(Dataset$cross)

Dataset$bushiness[Dataset$bushiness==0] <- NA

head(Dataset)

```

Create a dataset excluding control to run the survival model:

```{r}
Dataset_nc<-subset(Dataset, Dataset$trt != "control") #Dataset used for the model with survival as response variable
```


## Fitted models


```{r}

glmer_surv_nc <- glmmTMB(survived~ trt + colors + trt:colors + (1|block) +(1|cross), family = "binomial", data = 
                         Dataset_nc)

glmer_f <- glmmTMB(flowering~trt + colors + trt:colors + (1|block) +(1|cross), family = "binomial", data = Dataset)

glmer_nf_d <- glmmTMB(flower_count ~ trt + colors + trt:colors, data = Dataset, family = "nbinom2")


lmer_height<-lmer(height_cm~trt + colors + trt:colors + (1|block) +(1|cross), Dataset, na.action=na.exclude)
lmer_height.o<-lmer(height_cm~trt + colors + trt:colors + (1|block)+(1|cross), Dataset, na.action=na.omit) # This model only changes the na.action because packages used for diagnostics cant handle na.exclude() and some plotting functions handle na.omit() better. This does not affect any estimate and both models are equivalent as both na.exclude and na.omit remove the row where any of the dependent or independent variable is missing


glmer_bush <- glmmTMB(bushiness~ trt + colors + trt:colors + (1|block)+(1|cross), data = Dataset, family = "poisson")


glmer_tf <- glmmTMB(days_1flower~trt + colors + trt:colors + (1|block) +(1|cross), family = "nbinom2", data = Dataset)

```


## Model evaluations

```{r}

# Survival

simulationOutput_surv_d<-simulateResiduals(glmer_surv_nc) 
dharma_plot_surv<-plot(simulationOutput_surv_d)
surv_dharma<-testResiduals(simulationOutput_surv_d)


# Flowering

simulationOutput_flow<-simulateResiduals(glmer_f) 
dharma_plot_flow<-plot(simulationOutput_flow)
flow_dharma<-testResiduals(simulationOutput_flow) 


# flower_count


simulationOutput_nflow<-simulateResiduals(glmer_nf_d) 
dharma_plot_nflow<-plot(simulationOutput_nflow)



# Height

simulationOutput_height_d<-simulateResiduals(lmer_height.o)
dharma_plot_height<-plot(simulationOutput_height_d)
height_dharma<-testResiduals(simulationOutput_height_d)


# Bushiness

simulationOutput_bush_d<-simulateResiduals(glmer_bush)
dharma_plot_bush<-plot(simulationOutput_bush_d)
bush_dharma<-testResiduals(simulationOutput_bush_d) 


# t flower

simulationOutput_tflow_d<-simulateResiduals(glmer_tf)
dharma_plot_tflow<-plot(simulationOutput_tflow_d)
tflow_dharma<-testResiduals(simulationOutput_tflow_d)


```



## Summary Anova tables

```{r}

s_surv<-car::Anova(glmer_surv_nc,type="II")
s_flow<-car::Anova(glmer_f,type="II")
s_nflow<-car::Anova(glmer_nf_d,type="II")
s_height<-car::Anova(lmer_height,type="III")
s_bush<-car::Anova(glmer_bush,type="III")
s_tflow<-car::Anova(glmer_tf,type="III")
#s_iheight<-car::Anova(lm_iheight_col,type="III")

options(scipen = 999)

surv<-as.data.frame(s_surv)
colnames(surv)[1] <- "Chisq"
surv <- cbind(surv,rep("survival", nrow(surv)))
colnames(surv)[ncol(surv)] = "trait"

flow<-as.data.frame(s_flow)
colnames(flow)[1] <- "Chisq"
flow <- cbind(flow,rep("flowering", nrow(flow)))
colnames(flow)[ncol(flow)] = "trait"


nflow<-as.data.frame(s_nflow)
nflow <- cbind(nflow,rep("n flowers", nrow(nflow)))
colnames(nflow)[ncol(nflow)] = "trait"


height<-as.data.frame(s_height)
height <- cbind(height,rep("height", nrow(height)))
colnames(height)[ncol(height)] = "trait"


bush<-as.data.frame(s_bush)
bush <- cbind(bush,rep("bushiness", nrow(bush)))
colnames(bush)[ncol(bush)] = "trait"


tflow<-as.data.frame(s_tflow)
colnames(tflow)[1] <- "Chisq"
tflow <- cbind(tflow,rep("time to flower", nrow(tflow)))
colnames(tflow)[ncol(tflow)] = "trait"


#iheight<-as.data.frame(s_iheight)

df_anovas <- rbind(surv,
      flow,
      nflow,
      tflow,
      height,
      bush)

df_anovas$"Pr(>Chisq)"<- round(df_anovas$"Pr(>Chisq)", 4)
df_anovas$"Chisq"<- round(df_anovas$"Chisq", 2)

df_anovas

#write.csv(df_anovas, "../tables/df_anovas_traits.csv")

```





## Group comparisons of predicted means & plotting


Plotting settings

```{r}

my_theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5))

pd = position_dodge(0.9)
alpha_raw_points = 0.15

trt_order <- c("control", "early drought", "late drought")

Dataset$trt <- factor(Dataset$trt,
    levels = trt_order,ordered = TRUE)

```




### ----- Survival - all levels

Compute emmeans and CI

```{r}

marginal_surv1.1 = emmeans(glmer_surv_nc,
                   ~ colors:trt,  calc = c(n = ".wgt."))

back.emmeans_surv<-confint(marginal_surv1.1, type = "response")

```
Add letter display from tukey tests

```{r}


surv_emmeans_cld <- multcomp::cld(object = marginal_surv1.1,
                           Letters = letters)


surv_emmeans_cld <- surv_emmeans_cld %>% 
  arrange(colors) %>% 
  select(trt, .group, colors)


back.emmeans_surv <- left_join(back.emmeans_surv, surv_emmeans_cld, by = c("trt","colors"))

back.emmeans_surv

```



```{r}


Dataset_percentage <- aggregate(survived ~ trt + colors, data = Dataset, FUN = mean)

Dataset_percentage$trt <- factor(Dataset_percentage$trt,
   levels = trt_order,ordered = TRUE)



# Plotting raw data points
surv_raw <- ggdotchart(
  Dataset_percentage, 
  x = "trt", 
  y = "survived",
  color = "colors", 
  palette = c("black", "black", "black"), 
  alpha = 0.3,
  size = 5, 
  shape = 21,
  #add = "segment", 
  #add.params = list(color = "lightgray", size = 1.3),
  position = position_dodge(0.6),
  ggtheme = theme_classic(),
  sorting = "none",
  order = trt_order
)

# Overlay emmeans and CI on the raw data plot
surv_overlay_plot <- surv_raw +
  geom_errorbar(
    data = back.emmeans_surv,
    aes(
      x = trt, #factor(trt, levels = trt_order), 
      y = prob,  # Assuming prob is the probability in decimal form
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      color = colors
    ),
    position = position_dodge(width = 0.6),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_surv,
    aes(
      x = trt, #factor(trt, levels = trt_order), 
      y = prob,  # Assuming prob is the probability in decimal form
      color = colors
    ),
    position = position_dodge(width = 0.6),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
 #geom_text(
  #  data = back.emmeans_surv,
  #  aes(
  #    x = trt, # Adjust position based on factor levels,
  #    color = colors,
  #    y = asymp.UCL + 0.02,  # Adjust the position of the letters
 #     label = .group  # Use the .group column for labels
 #   ),
 #   position = position_dodge(width = 0.6),
 #   vjust = -0.5,  # Adjust vertical position
 #   size = 4
 # ) +
  scale_fill_manual(values = c("#990099", "#FF6666", "gold")) +
  scale_color_manual(
    values = c("#990099", "#FF6666", "gold"),  # Remove black color
    labels = c("pseudomajus", "F1 Hybrid", "striatum")  # Update labels
  ) +
  ylim(0,1.05)+
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  #labs(color = "Phenotype") +
  theme(legend.position = "none") +
  xlab(NULL) +
  ylab("survival probability") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))


surv_overlay_plot





```




### ----- Survival - significant levels

Compute emmeans and CI

```{r}

marginal_surv1.1 = emmeans(glmer_surv_nc,
                   ~ colors,  calc = c(n = ".wgt."))

back.emmeans_surv<-confint(marginal_surv1.1, type = "response")

```
Add letter display from tukey tests

```{r}


surv_emmeans_cld <- multcomp::cld(object = marginal_surv1.1,
                           Letters = letters)


surv_emmeans_cld <- surv_emmeans_cld %>% 
  arrange(colors) %>% 
  select(.group, colors)


back.emmeans_surv <- left_join(back.emmeans_surv, surv_emmeans_cld, by = c("colors"))

back.emmeans_surv

```



```{r}


Dataset_percentage <- aggregate(survived ~ colors, data = Dataset, FUN = mean)



# Plotting raw data points
surv_raw <- ggdotchart(
  Dataset_percentage, 
  x = "colors", 
  y = "survived",
  color = "colors", 
  palette = c("black", "black", "black"), 
  alpha = 0.3,
  size = 5, 
  shape = 21,
  #add = "segment", 
  #add.params = list(color = "lightgray", size = 1.3),
  position = position_dodge(0.6),
  ggtheme = theme_classic(),
  sorting = "none"
)

# Overlay emmeans and CI on the raw data plot
surv_overlay_plot2 <- surv_raw +
  geom_errorbar(
    data = back.emmeans_surv,
    aes(
      x = colors, #factor(trt, levels = trt_order), 
      y = prob,  # Assuming prob is the probability in decimal form
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      color = colors
    ),
    position = position_dodge(width = 0.6),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_surv,
    aes(
      x = colors, #factor(trt, levels = trt_order), 
      y = prob,  # Assuming prob is the probability in decimal form
      color = colors
    ),
    position = position_dodge(width = 0.6),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
 geom_text(
    data = back.emmeans_surv,
    aes(
      x = colors, # Adjust position based on factor levels,
      color = colors,
     y = asymp.UCL + 0.02,  # Adjust the position of the letters
      label = .group  # Use the .group column for labels
    ),
    position = position_dodge(width = 0.6),
   vjust = -0.5,  # Adjust vertical position
   size = 4
  ) +
  scale_fill_manual(values = c("#990099", "#FF6666", "gold")) +
  scale_color_manual(
    values = c("#990099", "#FF6666", "gold"),  # Remove black color
   # labels = c("pseudomajus", "F1 Hybrid", "striatum")  # Update labels
  ) +
  ylim(0,1.05)+
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  theme(legend.position = "none") +
  ylab("survival probability") +
  xlab(NULL)+
  theme(axis.title.y = element_text(margin = margin(r = 10)))+
   scale_x_discrete(guide = guide_axis(n.dodge=2),
                    labels = c("pseudomajus", "F1 Hybrid", "striatum"))


surv_overlay_plot2





```



### ----- Flowering - All levels


Compute emmeans and CI

```{r}
emmeans_flowering = emmeans(glmer_f,
                   ~ colors:trt,  calc = c(n = ".wgt."))

back.emmeans_flowering<-confint(emmeans_flowering, type = "response")

```

Add letter display from tukey tests

```{r}


flow_emmeans_cld <- multcomp::cld(object = emmeans_flowering,
                           Letters = letters)

flow_emmeans_cld <- flow_emmeans_cld %>% 
  arrange(trt, colors) %>% 
  select(.group, colors, trt)


back.emmeans_flowering <- left_join(back.emmeans_flowering, flow_emmeans_cld, by = c("trt", "colors"))

back.emmeans_flowering

```


```{r}

Dataset_percentage <- aggregate(flowering ~ trt + colors, data = Dataset, FUN = mean)


# Plotting raw data points
flow_raw <- ggdotchart(
  Dataset_percentage, 
  x = "trt", 
  y = "flowering",
  color = "colors", 
  palette = c("black", "black", "black"), 
  alpha = 0.3,
  size = 5, 
  shape = 21,
  #add = "segment", 
  #add.params = list(color = "lightgray", size = 1.3),
  position = position_dodge(0.6),
  ggtheme = theme_classic(),
    sorting = "none",
  order = trt_order
)

# Overlay emmeans and CI on the raw data plot
flow_overlay_plot <- flow_raw +
  geom_errorbar(
    data = back.emmeans_flowering,
    aes(
      x = trt, #factor(trt, levels = c(trt_order)), 
      y = prob,  # Assuming prob is the probability in decimal form
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      color = colors
    ),
    position = position_dodge(width = 0.6),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_flowering,
    aes(
      x = trt, #factor(trt, levels = c(trt_order)), 
      y = prob,  # Assuming prob is the probability in decimal form
      color = colors
    ),
    position = position_dodge(width = 0.6),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
 # geom_text(
 #   data = back.emmeans_flowering,
 #   aes(
 #     x = trt, # Adjust position based on factor levels,
  #    color = colors,
  #    y = asymp.UCL + 0.02,  # Adjust the position of the letters
  #    label = .group  # Use the .group column for labels
 #   ),
 #   position = position_dodge(width = 0.6),
 #   vjust = -0.5,  # Adjust vertical position
 #   size = 4
#  ) +
  scale_fill_manual(values = c("#990099", "#FF6666", "gold")) +
  scale_color_manual(
    values = c("#990099", "#FF6666", "gold"),  # Remove black color
    labels = c("pseudomajus", "F1 Hybrid", "striatum")  # Update labels
  ) +
  ylim(0, 1.05)+
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  #labs(color = "Phenotype") +
  theme(legend.position = "none")+
  xlab(NULL) +
  ylab("flowering probability") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))


flow_overlay_plot

```


### ----- Flowering - significant levels

#### colors

Compute emmeans and CI

```{r}
emmeans_flowering = emmeans(glmer_f,
                   ~ colors,  calc = c(n = ".wgt."))

back.emmeans_flowering<-confint(emmeans_flowering, type = "response")

```

Add letter display from tukey tests

```{r}


flow_emmeans_cld <- multcomp::cld(object = emmeans_flowering,
                           Letters = letters)

flow_emmeans_cld <- flow_emmeans_cld %>% 
  arrange(colors) %>% 
  select(.group, colors)


back.emmeans_flowering <- left_join(back.emmeans_flowering, flow_emmeans_cld, by = "colors")

back.emmeans_flowering

```


```{r}

Dataset_percentage <- aggregate(flowering ~ colors, data = Dataset, FUN = mean)


# Plotting raw data points
flow_raw <- ggdotchart(
  Dataset_percentage, 
  x = "colors", 
  y = "flowering",
  color = "colors", 
  palette = c("black", "black", "black"), 
  alpha = 0.3,
  size = 5, 
  shape = 21,
  #add = "segment", 
  #add.params = list(color = "lightgray", size = 1.3),
  position = position_dodge(0.6),
  ggtheme = theme_classic(),
    sorting = "none"
)

# Overlay emmeans and CI on the raw data plot
flow_overlay_plot2 <- flow_raw +
  geom_errorbar(
    data = back.emmeans_flowering,
    aes(
      x = colors, #factor(trt, levels = c(trt_order)), 
      y = prob,  # Assuming prob is the probability in decimal form
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      color = colors
    ),
    position = position_dodge(width = 0.6),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_flowering,
    aes(
      x = colors, #factor(trt, levels = c(trt_order)), 
      y = prob,  # Assuming prob is the probability in decimal form
      color = colors
    ),
    position = position_dodge(width = 0.6),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
  geom_text(
    data = back.emmeans_flowering,
    aes(
      x = colors, # Adjust position based on factor levels,
      color = colors,
      y = asymp.UCL + 0.02,  # Adjust the position of the letters
      label = .group  # Use the .group column for labels
    ),
    position = position_dodge(width = 0.6),
    vjust = -0.5,  # Adjust vertical position
    size = 4
 ) +
  scale_fill_manual(values = c("#990099", "#FF6666", "gold")) +
  scale_color_manual(
    values = c("#990099", "#FF6666", "gold"),  # Remove black color
    labels = c("pseudomajus", "F1 Hybrid", "striatum")  # Update labels
  ) +
  ylim(0, 1.05)+
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  #labs(color = "Phenotype") +
  theme(legend.position = "none") + 
  xlab(NULL) +
  ylab("flowering probability") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))+
   scale_x_discrete(guide = guide_axis(n.dodge=2),
                    labels = c("pseudomajus", "F1 Hybrid", "striatum"))


flow_overlay_plot2

```


#### trt

Compute emmeans and CI

```{r}
emmeans_flowering = emmeans(glmer_f,
                   ~ trt,  calc = c(n = ".wgt."))

back.emmeans_flowering<-confint(emmeans_flowering, type = "response")

```

Add letter display from tukey tests

```{r}


flow_emmeans_cld <- multcomp::cld(object = emmeans_flowering,
                           Letters = letters)

flow_emmeans_cld <- flow_emmeans_cld %>% 
  arrange(trt) %>% 
  select(.group, trt)


back.emmeans_flowering <- left_join(back.emmeans_flowering, flow_emmeans_cld, by = "trt")

back.emmeans_flowering

```


```{r}

Dataset_percentage <- aggregate(flowering ~ trt, data = Dataset, FUN = mean)


# Plotting raw data points
flow_raw <- ggdotchart(
  Dataset_percentage, 
  x = "trt", 
  y = "flowering",
  #color = "colors", 
  palette = c("black", "black", "black"), 
  alpha = 0.3,
  size = 5, 
  shape = 21,
  #add = "segment", 
  #add.params = list(color = "lightgray", size = 1.3),
  position = position_dodge(0.6),
  ggtheme = theme_classic(),
    sorting = "none"
)

# Overlay emmeans and CI on the raw data plot
flow_overlay_plot3 <- flow_raw +
  geom_errorbar(
    data = back.emmeans_flowering,
    aes(
      x = trt, #factor(trt, levels = c(trt_order)), 
      y = prob,  # Assuming prob is the probability in decimal form
      ymin = asymp.LCL,
      ymax = asymp.UCL,
    #  color = colors
    ),
    position = position_dodge(width = 0.6),
    width = 0.3,
    size = 1
  ) +
  geom_point(
    data = back.emmeans_flowering,
    aes(
      x = trt, #factor(trt, levels = c(trt_order)), 
      y = prob,  # Assuming prob is the probability in decimal form
   #   color = colors
    ),
    position = position_dodge(width = 0.6),
    size = 3,
    alpha = 0.5,
    shape = 15
  ) +
  geom_text(
    data = back.emmeans_flowering,
    aes(
      x = trt, # Adjust position based on factor levels,
     # color = colors,
      y = asymp.UCL + 0.02,  # Adjust the position of the letters
      label = .group  # Use the .group column for labels
    ),
    position = position_dodge(width = 0.6),
    vjust = -0.5,  # Adjust vertical position
    size = 4
 ) +
 # scale_fill_manual(values = c("#990099", "#FF6666", "gold")) +
 # scale_color_manual(
#    values = c("#990099", "#FF6666", "gold"),  # Remove black color
#    labels = c("pseudomajus", "F1 Hybrid", "striatum")  # Update labels
#  ) +
  ylim(0, 1.05)+
  theme_classic() +
  my_theme +
  guides(fill = FALSE) +
  labs(color = "Phenotype") +
  xlab(NULL) +
  ylab("flowering probability") +
  theme(axis.title.y = element_text(margin = margin(r = 10)))+
   scale_x_discrete(guide = guide_axis(n.dodge=2))


flow_overlay_plot3

```


### ----- N flowers - All levels

Compute emmeans and CI


```{r}
### N flowers (1 drought level) - treatment significant

marginal_nf.d = emmeans(glmer_nf_d,
                   ~ colors:trt,  calc = c(n = ".wgt."))

back.emmeans_nflow<-confint(marginal_nf.d, type = "response")


back.emmeans_nflow[which(is.infinite(back.emmeans_nflow$asymp.UCL)),"asymp.UCL"] <- 0

```



Add letter display from tukey tests

```{r}


nflow_emmeans_cld <- multcomp::cld(object = marginal_nf.d,
                           Letters = letters)

nflow_emmeans_cld <- nflow_emmeans_cld %>% 
  arrange(trt, colors) %>% 
  select(.group, colors, trt)


back.emmeans_nflow <- left_join(back.emmeans_nflow, nflow_emmeans_cld, by = c("trt", "colors"))

back.emmeans_nflow

```





```{r}


nflow_raw <- ggplot(Dataset, aes(x = trt, y = flower_count, group = colors)) +
  geom_jitter(
    aes(color = "black", fill = "black"),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  scale_fill_manual(values=c("black", "black", "black"))+
  scale_color_manual(values=c("black", "black", "black"))+
  theme_classic() +
  my_theme

  



# Overlay emmeans and CI on the raw data plot
nflow_overlay_plot <- nflow_raw +
  geom_errorbar(
    data = back.emmeans_nflow,
    aes(x = trt, y = response, ymin = asymp.LCL, ymax = asymp.UCL, color = colors),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  
  geom_point(
    data = back.emmeans_nflow,
    aes(x = trt, y = response, color = colors),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
    
  ) +
#  geom_text(
 #   data = back.emmeans_nflow,
 #   aes(
 #     x = trt, # Adjust position based on factor levels,
 #     color = colors,
 #     y = asymp.UCL + 0.1,  # Adjust the position of the letters
 #     label = .group  # Use the .group column for labels
  #  ),
 #   position = position_dodge(width = 0.75),
 #   vjust = -0.5,  # Adjust vertical position
 #   size = 4
#  ) +
    scale_fill_manual(values=c("#990099", "#FF6666", "gold"))+
  scale_color_manual(values=c("black", "#990099", "#FF6666", "gold"),labels = c("raw data", "pseudomajus", "F1 hybrid", "striatum"))+
  theme_classic()+
  my_theme+
guides(fill = FALSE)+
#labs(color = "phenotype")+
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("pseudomajus", "F1 Hybrid", "striatum")) + 
  xlab(NULL)+
  ylab("flower count")+
  theme(axis.title.y = element_text(margin = margin(r = 10)))
  

nflow_overlay_plot


```


### ----- N flowers - significant levels


#### colors

Compute emmeans and CI


```{r}
### N flowers (1 drought level) - treatment significant

marginal_nf.d = emmeans(glmer_nf_d,
                   ~ colors,  calc = c(n = ".wgt."))

back.emmeans_nflow<-confint(marginal_nf.d, type = "response")


back.emmeans_nflow[which(is.infinite(back.emmeans_nflow$asymp.UCL)),"asymp.UCL"] <- 0

```



Add letter display from tukey tests

```{r}


nflow_emmeans_cld <- multcomp::cld(object = marginal_nf.d,
                           Letters = letters)

nflow_emmeans_cld <- nflow_emmeans_cld %>% 
  arrange(colors) %>% 
  select(.group, colors)


back.emmeans_nflow <- left_join(back.emmeans_nflow, nflow_emmeans_cld, by =  "colors")

# we manually assign a different letter to hybrids because they have 0 flowers

back.emmeans_nflow[back.emmeans_nflow$colors == "M_Y", ".group"] <- "b"

back.emmeans_nflow

```





```{r}



# Overlay emmeans and CI on the raw data plot
nflow_overlay_plot2 <- ggplot(back.emmeans_nflow, aes(x = colors, y = response, group = colors)) +
  #nflow_raw +
  geom_errorbar(
    data = back.emmeans_nflow,
    aes(x = colors, y = response, ymin = asymp.LCL, ymax = asymp.UCL, color = colors),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  
  geom_point(
    data = back.emmeans_nflow,
    aes(x = colors, y = response, color = colors),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
    
  ) +
 geom_text(
    data = back.emmeans_nflow,
    aes(
      x = colors, # Adjust position based on factor levels,
      color = colors,
      y = asymp.UCL + 0.1,  # Adjust the position of the letters
      label = .group  # Use the .group column for labels
    ),
    position = position_dodge(width = 0.75),
    vjust = -0.5,  # Adjust vertical position
    size = 4
  ) +
    scale_fill_manual(values=c("#990099", "#FF6666", "gold"))+
  scale_color_manual(values=c("#990099", "#FF6666", "gold"),labels = c("pseudomajus", "F1 hybrid", "striatum"))+
  theme_classic()+
  my_theme+
guides(fill = FALSE)+
#labs(color = "phenotype")+
  theme(legend.position = "none") + 
  xlab(NULL)+
  ylab("flower count")+
  theme(axis.title.y = element_text(margin = margin(r = 10)))+
   scale_x_discrete(guide = guide_axis(n.dodge=2),
                    labels = c("pseudomajus", "F1 Hybrid", "striatum"))
  

nflow_overlay_plot2


```






#### trt

Compute emmeans and CI


```{r}
### N flowers (1 drought level) - treatment significant

marginal_nf.d = emmeans(glmer_nf_d,
                   ~ trt,  calc = c(n = ".wgt."))

back.emmeans_nflow<-confint(marginal_nf.d, type = "response")


back.emmeans_nflow[which(is.infinite(back.emmeans_nflow$asymp.UCL)),"asymp.UCL"] <- 0

```



Add letter display from tukey tests

```{r}


nflow_emmeans_cld <- multcomp::cld(object = marginal_nf.d,
                           Letters = letters)

nflow_emmeans_cld <- nflow_emmeans_cld %>% 
  arrange(trt) %>% 
  select(.group, trt)


back.emmeans_nflow <- left_join(back.emmeans_nflow, nflow_emmeans_cld, by =  "trt")

# manually assign different letter to late drought

back.emmeans_nflow[back.emmeans_nflow$trt == "late drought", ".group"] <- "c"

back.emmeans_nflow

```





```{r}



# Overlay emmeans and CI on the raw data plot
nflow_overlay_plot3 <- ggplot(back.emmeans_nflow, aes(x = trt, y = response)) +
  #nflow_raw +
  geom_errorbar(
    data = back.emmeans_nflow,
    aes(x = trt, y = response, ymin = asymp.LCL, ymax = asymp.UCL),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  
  geom_point(
    data = back.emmeans_nflow,
    aes(x = trt, y = response),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
    
  ) +
 geom_text(
    data = back.emmeans_nflow,
    aes(
      x = trt, # Adjust position based on factor levels,
      y = asymp.UCL + 0.1,  # Adjust the position of the letters
      label = .group  # Use the .group column for labels
    ),
    position = position_dodge(width = 0.75),
    vjust = -0.5,  # Adjust vertical position
    size = 4
  ) +
  theme_classic()+
  my_theme+
guides(fill = FALSE)+
labs(color = "phenotype")+
  xlab(NULL)+
  ylab("flower count")+
  theme(axis.title.y = element_text(margin = margin(r = 10)))+
   scale_x_discrete(guide = guide_axis(n.dodge=2))
  

nflow_overlay_plot3


```









### ----- Height


```{r}

emmeans_height = emmeans(lmer_height,
                   ~ colors:trt,  calc = c(n = ".wgt."))

emmeans_height <- multcomp::cld(object = emmeans_height,
                                   Letters = letters)

```




RAW DATA



```{r}


height_raw <- ggplot(Dataset, aes(x = trt, y = height_cm, group = colors)) +
  geom_jitter(
    aes(color = "black", fill = "black"),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  scale_fill_manual(values=c("black", "black", "black"))+
  scale_color_manual(values=c("black", "black", "black"))+
  theme_classic() +
  my_theme

  



# Overlay emmeans and CI on the raw data plot
height_overlay_plot <- height_raw +
  geom_errorbar(
    data = emmeans_height,
    aes(x = trt, y = emmean, ymin = lower.CL, ymax = upper.CL, color = colors),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  
  geom_point(
    data = emmeans_height,
    aes(x = trt, y = emmean, color = colors),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
    
  ) +
  geom_text(
    data = emmeans_height,
    aes(
      x = trt, # Adjust position based on factor levels,
      color = colors,
      y = upper.CL + 0.1,  # Adjust the position of the letters
      label = .group  # Use the .group column for labels
    ),
    position = position_dodge(width = 0.75),
    vjust = -0.5,  # Adjust vertical position
    size = 4
  ) +
    scale_fill_manual(values=c("#990099", "#FF6666", "gold"))+
  scale_color_manual(values=c("black", "#990099", "#FF6666", "gold"),labels = c("raw data", "pseudomajus", "F1 hybrid", "striatum"))+
  theme_classic()+
  my_theme+
guides(fill = FALSE)+
labs(color = "phenotype")+
  xlab(NULL)+
  ylab("height (cm)")+
  theme(axis.title.y = element_text(margin = margin(r = 10)))
  

height_overlay_plot


```



### ----- Bushiness

Compute emmeans and CI

```{r}
emmeans_bush = emmeans(glmer_bush,
                   ~ colors:trt,  calc = c(n = ".wgt."))

back.emmeans_bush<-confint(emmeans_bush, type = "response")


```


Add letter display from tukey tests

```{r}

emmeans_bush_cld <- multcomp::cld(object = emmeans_bush,
                                   Letters = letters)

emmeans_bush_cld <- emmeans_bush_cld %>% 
  arrange(trt, colors) %>% 
  select(.group, colors, trt)


back.emmeans_bush <- left_join(back.emmeans_bush, emmeans_bush_cld, by = c("trt", "colors"))

back.emmeans_bush

```




```{r}


bush_raw <- ggplot(Dataset, aes(x = trt, y = bushiness, group = colors)) +
  geom_jitter(
    aes(color = "black", fill = "black"),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  scale_fill_manual(values=c("black", "black", "black"))+
  scale_color_manual(values=c("black", "black", "black"))+
  theme_classic() +
  my_theme

  



# Overlay emmeans and CI on the raw data plot
bushiness_overlay_plot <- bush_raw +
  geom_errorbar(
    data = back.emmeans_bush,
    aes(x = trt, y = rate, ymin = asymp.LCL, ymax = asymp.UCL, color = colors),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  
  geom_point(
    data = back.emmeans_bush,
    aes(x = trt, y = rate, color = colors),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
    
  ) +
  geom_text(
    data = back.emmeans_bush,
    aes(
      x = trt, # Adjust position based on factor levels,
      color = colors,
      y = asymp.UCL + 0.1,  # Adjust the position of the letters
      label = .group  # Use the .group column for labels
    ),
    position = position_dodge(width = 0.75),
    vjust = -0.5,  # Adjust vertical position
    size = 4
  ) +
    scale_fill_manual(values=c("#990099", "#FF6666", "gold"))+
  scale_color_manual(values=c("black", "#990099", "#FF6666", "gold"),labels = c("raw data", "pseudomajus", "F1 hybrid", "striatum"))+
  theme_classic()+
  my_theme+
guides(fill = FALSE)+
labs(color = "phenotype")+
  xlab(NULL)+
  ylab("bushiness (n nodes)")+
  theme(axis.title.y = element_text(margin = margin(r = 10)))
  

bushiness_overlay_plot


```





### ----- Time to flower


Compute emmeans and CI

```{r}
emmeans_tflow = emmeans(glmer_tf,
                   ~ colors:trt,  calc = c(n = ".wgt."))

back.emmeans_tflow<-confint(emmeans_tflow, type = "response")

```


Add letter display from tukey tests

```{r}

emmeans_tflow_cld <- multcomp::cld(object = emmeans_tflow,
                                   Letters = letters)

emmeans_tflow_cld <- emmeans_tflow_cld %>% 
  arrange(trt, colors) %>% 
  select(.group, colors, trt)


back.emmeans_tflow <- left_join(back.emmeans_tflow, emmeans_tflow_cld, by = c("trt", "colors"))

back.emmeans_tflow

```






```{r}
tflow_raw <- ggplot(Dataset, aes(x = trt, y = days_1flower, group = colors)) +
  geom_jitter(
    aes(color = "black", fill = "black"),
    position = position_jitterdodge(0.08),
    size = 1.2,
    alpha = alpha_raw_points
  ) +
  scale_fill_manual(values=c("black", "black", "black"))+
  scale_color_manual(values=c("black", "black", "black"))+
  theme_classic() +
  my_theme

  



# Overlay emmeans and CI on the raw data plot
tflow_overlay_plot <- tflow_raw +
  geom_errorbar(
    data = back.emmeans_tflow,
    aes(x = trt, y = response, ymin = asymp.LCL, ymax = asymp.UCL, color = colors),
    position = position_dodge(width = 0.75),
    width = 0.3,
    size = 1
  ) +
  
  geom_point(
    data = back.emmeans_tflow,
    aes(x = trt, y = response, color = colors),
    position = position_dodge(width = 0.75),
    size = 3,
    alpha = 0.5,
    shape = 15
    
  )  +
  geom_text(
    data = back.emmeans_tflow,
    aes(
      x = trt, # Adjust position based on factor levels,
      color = colors,
      y = asymp.UCL + 0.1,  # Adjust the position of the letters
      label = .group  # Use the .group column for labels
    ),
    position = position_dodge(width = 0.75),
    vjust = -0.5,  # Adjust vertical position
    size = 4
  ) +
    scale_fill_manual(values=c("#990099", "#FF6666", "gold"))+
  scale_color_manual(values=c("black", "#990099", "#FF6666", "gold"),labels = c("raw data", "pseudomajus", "F1 hybrid", "striatum"))+
  theme_classic()+
  my_theme+
guides(fill = FALSE)+
labs(color = "phenotype")+
  ylim(38,95)+
  xlab(NULL)+
  ylab("time to flower (days)")+
  theme(axis.title.y = element_text(margin = margin(r = 10)))

tflow_overlay_plot

```




# Fig 1

```{r}

arranged_flowering_coltrt <- ggarrange(
  
  flow_overlay_plot2 + ylab(NULL), flow_overlay_plot3 + ylab(NULL),
  
  ncol = 2,
  nrow = 1,
  
  labels = c("D", "E")
)


arranged_nflow_coltrt <- ggarrange(
  
  nflow_overlay_plot2 + ylab(NULL) + ylim(0,1.8), nflow_overlay_plot3 + ylab(NULL) +ylim(0,6.5),
  
  ncol = 2,
  nrow = 1,
  
  labels = c("G", "H")
)



figure1 <- ggarrange(
  
  surv_overlay_plot +
  theme(axis.title.y = element_text(margin = margin(r = 10)))+
   scale_x_discrete(guide = guide_axis(n.dodge=2)), 
  
  surv_overlay_plot2 + ylab(NULL) + plot_spacer(),
  
  flow_overlay_plot+
  theme(axis.title.y = element_text(margin = margin(r = 10)))+
   scale_x_discrete(guide = guide_axis(n.dodge=2)), 
  
  arranged_flowering_coltrt,
  
  nflow_overlay_plot+
  theme(axis.title.y = element_text(margin = margin(r = 10)))+
   scale_x_discrete(guide = guide_axis(n.dodge=2)), 
 
   arranged_nflow_coltrt,
  
  nrow = 3,
  ncol = 2,
  
  common.legend = TRUE,
  
  labels = c("A", "B", "C", "", "F", "")
  
)

figure1

ggsave("../figures/figure1_test.png", height = 9, width = 10)


```




```{r}

figure_2_a <- ggarrange(
  
    surv_overlay_plot,

flow_overlay_plot,

nflow_overlay_plot,

nrow = 3,
ncol = 1,

common.legend = TRUE,

labels = c("A", "B", "C")
  
)


figure_2_a


#ggsave("../figures/figure_1.png", height = 8, width = 5)

```




```{r}


figure_2_b <- ggarrange(
  
    height_overlay_plot,

bushiness_overlay_plot,

tflow_overlay_plot,

nrow = 3,
ncol = 1,

common.legend = TRUE,

labels = c("A", "B", "C")
  
)

figure_2_b


#ggsave("../figures/figure_s1.png", height = 8, width = 5)

```




## save emmeans results

```{r}

options(scipen = 999)

colmnames_anova <- c("colors", "trt", "estimated response", "SE", "df", "n",
                     "lower CI", "upper CI", "comp" ,"trait")

df_anova_surv <- as.data.frame(back.emmeans_surv)
df_anova_surv$trait <- rep("survival", nrow(df_anova_surv))
colnames(df_anova_surv) = colmnames_anova

df_anova_flow <- as.data.frame(back.emmeans_flowering)
df_anova_flow$trait <- rep("flowering", nrow(df_anova_flow))
colnames(df_anova_flow) = colmnames_anova

df_anova_nflow <- as.data.frame(back.emmeans_nflow)
df_anova_nflow$trait <- rep("n flowers", nrow(df_anova_nflow))
colnames(df_anova_nflow) = colmnames_anova
 
df_anova_height <- as.data.frame(emmeans_height)
df_anova_height$trait <- rep("height (cm)", nrow(df_anova_height))
colnames(df_anova_height) = colmnames_anova

df_anova_bush <- as.data.frame(back.emmeans_bush)
df_anova_bush$trait <- rep("bushiness (n nodes)", nrow(df_anova_bush))
colnames(df_anova_bush) = colmnames_anova

df_anova_tflow <- as.data.frame(back.emmeans_tflow)
df_anova_tflow$trait <- rep("time to flower (days)", nrow(df_anova_tflow))
colnames(df_anova_tflow) = colmnames_anova

df_anovas <- do.call("rbind", list(df_anova_surv, 
                      df_anova_flow, 
                      df_anova_nflow,
                      df_anova_height,
                      df_anova_bush,
                      df_anova_tflow))

df_anovas$`estimated response` <- round(df_anovas$`estimated response`, 2)
df_anovas$`lower CI` <- round(df_anovas$`lower CI`, 2)
df_anovas$`upper CI` <- round(df_anovas$`upper CI`, 2)

df_anovas

#write.csv(df_anovas, "../tables/df_emmeans.csv")


```






























