---
title: "Risk death for flowering individuals"
output: html_notebook
---


This script analyses the risk of death of individuals based on wether they flower or not, for both early and late drought treatment.


“Risk” refers to the probability of occurrence of an event or outcome. Statistically, risk = chance of the outcome of interest/all possible outcomes.



Load packages

```{r}
library(ggplot2)
library(ggpubr)
library(reshape2)
library(tidyverse)

```


Read data

```{r}

Dataset <- read.csv("../dataset.csv")


Dataset <- Dataset %>% 
  filter(w_b == "b",
         trt != "control") 

head(Dataset)

```


Parameters for plotting

```{r}
my.theme<-theme(axis.text=element_text(size=12),
        axis.title = element_text(size = 14),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)))

```


Compute risk death:

```{r}

df_risk_death <- Dataset %>%
    group_by(trt, colors) %>%
    summarise(
      died_flowering = sum(survived == 0 & flowering == 1),
      survived_flowering = sum(survived == 1 & flowering == 1),
      died_non_flowering = sum(survived == 0 & flowering == 0),
      survived_non_flowering = sum(survived == 1 & flowering == 0)
    ) %>% 
  mutate(
    death_risk_flowering = died_flowering / (died_flowering + survived_flowering)*100,
    death_risk_non_flowering = died_non_flowering / (died_non_flowering + survived_non_flowering) * 100
  )

df_risk_death


```




## test for significant differences between flowering and non-flowering groups

We perform a two-groups proportion Z test to determine whether flowering vs non-flowering individuals of a given phenotype in a given treatment show significant differences in their % risk death.

* In this test, the null hypothesis is that there exist no differences between the groups.

Function to compute the 2 groups proportion test

```{r}

df_risk_death <- df_risk_death %>% 
  rowwise() %>%
  mutate(statistic = prop.test(x = c(died_flowering, died_non_flowering),
                               n = c((died_flowering+survived_flowering), (died_non_flowering + survived_non_flowering)))$statistic,
         
         estimate = paste(round(prop.test(x = c(died_flowering, died_non_flowering),
                               n = c((died_flowering+survived_flowering), (died_non_flowering + survived_non_flowering)))$estimate, digits = 2), collapse = " , "),
         
         pvalue = prop.test(x = c(died_flowering, died_non_flowering),
                               n = c((died_flowering+survived_flowering), (died_non_flowering + survived_non_flowering)))$p.value)


df_risk_death

#write.csv(df_risk_death, "../tables/df_risk_death.csv")

```




## Plot results

```{r}


death_risk_long <- df_risk_death %>%
  select(trt, colors, death_risk_flowering, death_risk_non_flowering) %>%
  pivot_longer(cols = c(death_risk_flowering, death_risk_non_flowering),
               names_to = "flowering",
               values_to = "death_risk") %>%
  mutate(flowering = if_else(flowering == "death_risk_flowering", "flowering", "non-flowering"))


# Plot using ggdotchart with facet_wrap
ggplot(death_risk_long, aes(x = flowering, y = death_risk, fill = colors, color = colors)) +
  geom_dotplot(binaxis = "y", stackdir = "center", position = position_dodge(width = 0.4), binwidth = 3.5) +
  facet_wrap(~ trt, scales = "free_x", nrow = 1) +
  scale_fill_manual(values = c("M_Y" = "#FF6666", "M_M" = "#990099", "Y_Y" = "gold")) +
  scale_color_manual(values = c("M_Y" = "#FF6666", "M_M" = "#990099", "Y_Y" = "gold")) +
  labs(x = NULL, y = "% Death Risk", fill = "Colors", color = "Colors") +  # Remove x-axis label and update color legends
  theme_classic() +
  theme(strip.text = element_text(size = 12))+
  my.theme

ggsave("../figures/death_risk.png", height = 4, width = 6)


```


```{r}
# Filter the dataframe for hybrids (M_Y) under late drought
df_subset <- df_risk_death %>%
  filter(trt == "late drought", colors == "M_Y")

# Extract the death risks
risk_flowering <- df_subset$death_risk_flowering
risk_non_flowering <- df_subset$death_risk_non_flowering

# Compute the risk ratio
risk_ratio <- risk_flowering / risk_non_flowering
risk_ratio

```































